---
title: "odi-batting"
author: "sambit s misra"
date: "4 February 2018"
output: html_document
---
# importing the dataset
```{r}
odi = read.csv("C:/Users/Administrator/Documents/sambit files/odi-batting.csv",header=T)
```

# checking data type of variables
```{r}
str(odi)
summary(odi)
```

# importing the libraries
```{r}
library(knitr)
library(dplyr)
library(lubridate)
```

# removing na values
```{r}
odi = na.omit(odi)
```

# top 10 country with maximum players 
```{r}
country_players = odi %>% group_by(Country) %>% summarise(totalplayers=n_distinct(Player)) %>% arrange(-totalplayers)
kable(country_players[1:10,])
```

# we see that england has the highest number of player records followed by india ,australia etc. intersting insight is that zimbabwe and bangladesh have more player records than south africa.

# top 10 country wise runs scored
```{r}

country_wise_runs = odi %>% group_by(Country) %>% summarise(totalruns = sum(Runs)) %>% arrange(-totalruns)
kable(country_wise_runs[1:10,])
```

# we see that india has the highest no of runs scored by its players followed closely by australia. within the test playing nations there is a significant decrease in runs scored after pakistan and england. zimbabwe and bangladesh are way below the rest.

# top 10 country with highest average score rate
```{r}
country_wise_scorerate = odi %>% group_by(Country) %>% summarise(mean_scorerate = mean(ScoreRate)) %>% arrange(-mean_scorerate)
kable(country_wise_scorerate[1:10,])
```

# there is an interesting insight here. south africa which was in the bottom five in runs scored is in the top of list in average score rate of its players. india which was in top of list in runs scored is in 4th place in this list. this means that indian players tend to play less aggressively than others. they play safe to score more runs. interestingly afghanistan is in the list.

# top 10 country with maximum number of matches played in 2010
```{r}
odi$MatchDate = as.Date(odi$MatchDate,format="%m-%d-%Y")
country_match = odi %>% filter(year(MatchDate)==2010)%>% group_by(Country) %>% summarise(totalmatch=n_distinct(MatchDate)) %>% arrange(-totalmatch)
kable(country_match[1:10,])
```

# bangladesh and india have played the maximum number of matches in 2010. interstingly zimbabwe has played more number of matches than england,pakistan and west indies in 2010. also ireland surprisingly is in the list.

# top 10 country with maximum number of matches played in lords
```{r}
country_lords = odi %>% filter(Ground=="Lord's") %>% group_by(Country) %>% summarise(lords_match=n_distinct(MatchDate)) %>% arrange(-lords_match)
kable(country_lords[,])
```

# as expected home country england has played the highest number of matches in lord's. india has played 7 matches in lord's including the famous natwest cup final win over england in 2002.

# top 10 country with max no of matches played against india
```{r}
country_against_india = odi %>% filter(Versus=="India") %>% group_by(Country) %>% summarise(against_india=n_distinct(MatchDate)) %>% arrange(-against_india)
kable(country_against_india[1:10,])
```

# we see asian countries like srilanka and pakistan have played more number of matches against india than other countries have played against india. a possible explanation to it can be the revenue they get because of a series against india.

# top 10 players who scored maximum runs
```{r}
player_runs = odi %>% group_by(Player) %>% summarise(totalruns=sum(Runs)) %>% arrange(-totalruns)
kable(player_runs[1:10,])
```

# we see that sachin is way ahead of others in total runs scored in odi. ganguly and dravid are the other indians  there in top 10 scorers list.

# top 10 players with highest average score rate
```{r}
player_scorerate = odi %>% group_by(Player) %>% summarise(avg_scorerate = mean(ScoreRate)) %>% arrange(-avg_scorerate)
kable(player_scorerate[1:10,])
```

# scorerate defines how much runs the player scored in 100 balls faced. it tells about how aggressive the batsman is. we can see no indian batsman in the list. this shows that indian batsman are rarely aggressive batsmen. they play more of a stable innings.

# top 10 players who played maximum matches in 2010
```{r}
player_match_2010 = odi %>% filter(year(MatchDate)==2010) %>% group_by(Player) %>% summarise(totalmatch=n()) %>% arrange(-totalmatch)
kable(player_match_2010[1:10,])
```

# virat kohli played 24 matches in 2010. but india played 27 matches in 2010. this means that kohli missed 3 matches due to injury or probably dropped from squad. bangladesh played 27 matches in 2010 and imrul and shakib played all the matches.

# top 10 players who played maximum matches in lord's
```{r}
player_lords = odi %>% filter(Ground=="Lord's") %>% group_by(Player) %>% summarise(lordsmatch=n()) %>% arrange(-lordsmatch)
kable(player_lords[1:10,])
```

# as expected all the 10 players are english players who played maximum number of matches in  lord's.

# top 10 players who played most matches against india
```{r}
player_against_india = odi %>% filter(Versus=="India") %>% group_by(Player) %>% summarise(totalmatch=n()) %>% arrange(-totalmatch)
kable(player_against_india[1:10,])
```

# most players in this list are asian players who played most number of matches against india. ricky ponting is the only exception in this list. 

# top 10 players of each country by runs
```{r}
country_top_scorer = odi %>% group_by(Country,Player) %>% summarise(totalruns=sum(Runs)) %>% arrange(-totalruns) %>% top_n(10) %>% arrange(Country)
```

# this gives the top 10 players of each country by total runs scored by them. 
